% TEMPLATE for Usenix papers, specifically to meet requirements of
%  USENIX '05
% originally a template for producing IEEE-format articles using LaTeX.
%   written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
% adapted by David Beazley for his excellent SWIG paper in Proceedings,
%   Tcl 96
% turned into a smartass generic template by De Clarke, with thanks to
%   both the above pioneers
% use at your own risk.  Complaints to /dev/null.
% make it two column with no page numbering, default is 10 point

% Munged by Fred Douglis <douglis@research.att.com> 10/97 to separate
% the .sty file from the LaTeX source template, so that people can
% more easily include the .sty file into an existing document.  Also
% changed to more closely follow the style guidelines as represented
% by the Word sample file. 

% Note that since 2010, USENIX does not require endnotes. If you want
% foot of page notes, don't include the endnotes package in the 
% usepackage command, below.

% This version uses the latex2e styles, not the very ancient 2.09 stuff.
\documentclass[letterpaper,twocolumn,10pt]{article}
%\usepackage{usenix,epsfig,endnotes}
\usepackage{usenix}
\usepackage{epsfig}
\usepackage{comment}
%\usepackage{subfigure}
\usepackage[pdftex]{color}
\usepackage{url}
\usepackage{listings}
\usepackage{textcomp}
\usepackage{gnuplot-lua-tikz}
\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage{minted}
%\usepackage{bm}
\usepackage{paralist}
%\usepackage{ulem}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}

\lstset{
	backgroundcolor=\color{lbcolor},
		tabsize=4,
		rulecolor=,
		language=matlab,
		basicstyle=\scriptsize,
		upquote=true,
		aboveskip={1.5\baselineskip},
		columns=fixed,
		showstringspaces=false,
		extendedchars=true,
		breaklines=true,
		prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
		frame=single,
		showtabs=false,
		showspaces=false,
		showstringspaces=false,
		identifierstyle=\ttfamily,
		keywordstyle=\color[rgb]{0,0,1},
		commentstyle=\color[rgb]{0.133,0.545,0.133},
		stringstyle=\color[rgb]{0.627,0.126,0.941},
}


\newcommand{\rb}[1]{{\color{red}{\it RB: #1}}}
\newcommand{\dz}[1]{{\color{blue}{\it DZ: #1}}}
\newcommand{\dm}[1]{{\color{green}{\it DM: #1}}}
\newcommand{\jv}[1]{{\color{orange}{\it JV: #1}}}

\newlength{\saveparindent}
\setlength{\saveparindent}{\parindent}
\newlength{\saveparskip}
\setlength{\saveparskip}{\parskip}

\newenvironment{newenumerate}{%
\begin{list}{({\rm \arabic{ctr}})\hfill}{\usecounter{ctr}\labelwidth=17pt%
\labelsep=6pt \leftmargin=23pt \topsep=.5pt%
\setlength{\listparindent}{\saveparindent}%
\setlength{\parsep}{\saveparskip}%
\setlength{\itemsep}{5pt} }}{\end{list}}

\newenvironment{newitemize}{%
\begin{list}{\hspace{3pt}$\bullet$\hfill}{\labelwidth=12pt%
\labelsep=3pt \leftmargin=15pt \topsep=1pt%
\setlength{\listparindent}{\saveparindent}%
\setlength{\parsep}{\saveparskip}%
\setlength{\itemsep}{1pt}}}{\end{list}}

\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{An SSD-based eigensolver for spectral analysis on billion-node graphs}

%for single author (just remove % characters)
\author{
{\rm Da Zheng, Randal Burns}\\
Department of Computer Science \\
Johns Hopkins University
\and
{\rm Joshua Vogelstein}\\
Institute for Computational Medicine \\
Department of Biomedical Engineering \\
Johns Hopkins University
% copy the following lines to add more authors
\and
{\rm Carey E. Priebe}\\
Department of Applied Mathematics and Statistics \\
Johns Hopkins University
\and
{\rm Alexander S. Szalay}\\
Department of Physics and Astronomy \\
Johns Hopkins University
} % end author

\maketitle

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
\thispagestyle{empty}


\subsection*{Abstract}
Many eigensolvers such as ARPACK and Anasazi have been developed to compute
eigenvalues of a large sparse matrix. These eigensolvers are limited by
the capacity of RAM. They run in memory of a single machine for smaller
eigenvalue problems and require the distributed memory for larger problems.

In contrast, we develop an SSD-based eigensolver framework called FlashEigen,
which extends Anasazi eigensolvers to SSDs, to compute eigenvalues of a graph
with hundreds of millions or even billions of vertices in a single machine.
FlashEigen performs sparse matrix multiplication in a semi-external memory
fashion, i.e., we keep the sparse matrix on SSDs and the dense matrix in memory.
%We specifically optimize the
%SpMM implementation for many real-world graphs that have nearly random
%vertex connection and power-law distribution in vertex degree. Due to
%the sparsity of the real-world graphs, the eigensolvers require large
%storage space for the vector subspace.
We store the entire vector subspace on SSDs and reduce I/O to improve
performance through caching the most recent dense matrix.
Our result shows that FlashEigen is able to achieve 40\%-60\% performance
of its in-memory implementation and has performance comparable to the Anasazi
eigensolvers on a machine with 48 CPU cores. Furthermore, it is capable of
scaling to a graph with 3.4 billion vertices and 129 billion edges. It takes
about four hours to compute eight eigenvalues of the billion-node graph using
120 GB memory.

\section{Introduction}
\input{intro}

\section{Related Work}
\input{relwork}

\input{design}

\input{eval}

\section{Conclusions}
\input{conclusion}

%\section{Acknowledgments}


{\footnotesize \bibliographystyle{acm}
\bibliography{eigensolver}  % sigproc.bib is the name of the Bibliography in this case


%\theendnotes

\end{document}
